Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-08 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Add figures to Madrzykowski thesis.

Figures 3 through 5 are graphs showing measurement results of heat release rate versus time for the natural gas, gasoline and polyurethane foam fueled fires. In each graph, the replicate heat release time histories are overlaid to give a sense of the experimental repeatability for each fuel.

@@

The natural gas burner provided the most repeatable results, with an average steady state heat release rate of approximately 80~kW $\pm$ 9~kW with an average total heat release of approximately 23.3~MJ $\pm$ 1.6~MJ over a 300~s period.  The gasoline, based on the average over a 30 s interval bounding the peak heat release rate of each of the 17 experiments, provided an average peak heat release rate of approximately 80~kW $\pm$ 24~kW.  The total heat release of the gasoline over the 300~s period after ignition was approximately 13.6~MJ $\pm$ 0.6~MJ.  The polyurethane foam demonstrated the largest variability, as shown in Figure 5, of the three fuels in terms of peak heat release rate and time to peak heat release.  Averaging over a 15~s interval bounding the peak heat release rate from each of the ten experiments yielded an average peak heat release rate of approximately 45~kW $\pm$ 21~kW.  The total heat release over the 300~s period after ignition was of 4.1~MJ $\pm$ 0.2~MJ

@@

Plots of the centerline temperatures averaged over the replicate experiments for each of the fuels are presented in Figures 6 through 8.  The number following “TC” in the legend of the graphs represents the height of the thermocouple in increments of 10~cm (or dm) above the fuel surface.  The natural gas temperatures were averaged over a ``steady-state'' burning period as reflected by the data in Figure 6.  The temperature differences between the vertical thermocouple locations, for locations greater than 20 cm above the fire location, begin to decrease as the distances from the top of the burner increase.  This trend is similar for both the gasoline fires and the polyurethane foam fires as well.  The most notable difference between the natural gas and the other fuels is the growth and decay of the fire during the test period.

@@

The repeatability of the total heat flux values varied in a manner similar to the heat release rates, in that the natural gas fires provided a steady state value while the gasoline and the polyurethane foam fires generated a fire curve that has periods of growth and decay.   The sensing surface of the heat flux gauges were located 0.95~m horizontally from the center of the burner/fuel and 0.50~m above the top of the burner/pan lip/fuel surface.  The average peak total heat flux for the natural gas fires were approximately 2.0~kW/m$^2$, as shown in the example data set in Figure 9.  The average peak values for the gasoline and polyurethane foam were approximately 3.5~kW/m$^2$ and 2.5~kW/m$^2$ respectively.  The values are only presented for the experiments which had two heat flux gauges installed as described in the instrumentation section above.

@@

Mass loss measurements were not possible for the natural gas burner experiments.  The mass loss versus time curves for the gasoline and the polyurethane experiments are shown in Figures 12 and 13.  The gasoline exhibits a linear mass loss rate for the initial 150~s of approximately 0.12~kg per minute.  The polyurethane foam was ignited with a small flame in the center of the top surface of the foam block.  As the flame spread across the surface it also began to burn down into the foam sample.  As a result the exposed surface of the foam began to take the shape of a bowl as the burning rate became linear at approximately 0.12~kg per minute.

@@

Examples of the photographic images are presented in Figures 14 and 15.  Figure 14 is a photograph of the natural gas flame taken during a period of ``steady state'' burning at approximately 80~kW and a photograph of the polyurethane foam fire taken near its peak heat release rate of approximately 60~kW.  Figure 9 presents two photographs of the same gasoline fueled fire when it was burning near its peak heat release rate of 100~kW.  The two photographs, taken 0.2~s apart, represent the range of the observed flame heights from one of the gasoline experiments from 0.65~m and 1.0~m, respectively.

>>>>

Figure~\ref{HRR} shows the measured heat release rates for the natural gas, gasoline and polyurethane foam fueled fires. In each graph, the replicate heat release time histories are overlaid to give a sense of the experimental repeatability for each fuel.

\begin{figure}[p]
  \centering
  \includegraphics[width=4.5in]{../Figures/Fig3}\\
  \includegraphics[width=4.5in]{../Figures/Fig4}\\
  \includegraphics[width=4.5in]{../Figures/Fig5}\\
  \caption[Heat release rates for the natural gas, gasoline and foam fires]{Heat release rates for the natural gas, gasoline, and foam fires.}
  \label{HRR}
\end{figure}

@@

The natural gas burner provided the most repeatable results, with an average steady state heat release rate of 80~kW $\pm$ 9~kW with an average total heat release of 23.3~MJ $\pm$ 1.6~MJ over a 300~s period.  The gasoline, based on the average over a 30 s interval bounding the peak heat release rate of each of the 17 experiments, provided an average peak heat release rate of 80~kW $\pm$ 24~kW.  The total heat release of the gasoline over the 300~s period after ignition was 13.6~MJ $\pm$ 0.6~MJ.  The polyurethane foam demonstrated the largest variability of the three fuels in terms of peak heat release rate and time to peak heat release.  Averaging over a 15~s interval bounding the peak heat release rate from each of the ten experiments yielded an average peak heat release rate of 45~kW $\pm$ 21~kW.  The total heat release over the 300~s period after ignition was 4.1~MJ $\pm$ 0.2~MJ

@@

Plots of the centerline temperatures averaged over the replicate experiments for each of the fuels are presented in Fig.~\ref{Temp}. The natural gas temperatures were averaged over a ``steady-state'' burning period.  The temperature differences between the vertical thermocouple locations, for locations greater than 20~cm above the fire location, begin to decrease as the distances from the top of the burner increase.  This trend is similar for both the gasoline fires and the polyurethane foam fires as well.  The most notable difference between the natural gas and the other fuels is the growth and decay of the fire during the test period.

\begin{figure}[p]
  \centering
  \includegraphics[width=4.5in]{../Figures/Fig6}\\
  \includegraphics[width=4.5in]{../Figures/Fig7}\\
  \includegraphics[width=4.5in]{../Figures/Fig8}\\
  \caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperature for the natural gas, gasoline, and foam fires. The number following the TC represents the height of thermocouple in decimeters  above the surface of the burner.}
  \label{Temp}
\end{figure}

@@

Plots of the total heat flux for each of the fuels are presented in Fig.~\ref{Flux}. The repeatability of the total heat flux values varied in a manner similar to the heat release rates, in that the natural gas fires provided a steady state value while the gasoline and the polyurethane foam fires generated a fire curve that has periods of growth and decay.   The sensing surface of the heat flux gauges were located 0.95~m horizontally from the center of the burner/fuel and 0.50~m above the top of the burner/pan lip/fuel surface.  The average peak total heat flux for the natural gas fires were approximately 2.0~kW/m$^2$.  The average peak values for the gasoline and polyurethane foam were approximately 3.5~kW/m$^2$ and 2.5~kW/m$^2$ respectively.  The values are only presented for the experiments which had two heat flux gauges installed as described in the instrumentation section above.

\begin{figure}[p]
  \centering
  \includegraphics[width=4.5in]{../Figures/Fig9}\\
  \includegraphics[width=4.5in]{../Figures/Fig10}\\
  \includegraphics[width=4.5in]{../Figures/Fig11}\\
  \caption[Total heat flux versus time for natural gas, gasoline, and foam fires]{Total heat flux versus time for natural gas, gasoline, and foam fires.}
  \label{Flux}
\end{figure}

@@

Mass loss measurements were not possible for the natural gas burner experiments.  The mass loss versus time curves for the gasoline and the polyurethane experiments are shown in Fig.~\ref{Mass}.  The gasoline exhibits a linear mass loss rate for the initial 150~s of approximately 2~g/s.  The polyurethane foam was ignited with a small flame in the center of the top surface of the foam block.  As the flame spread across the surface it also began to burn down into the foam sample.  As a result the exposed surface of the foam began to take the shape of a bowl as the burning rate became linear at approximately 2~g/s.

\begin{figure}[p]
  \centering
  \includegraphics[width=4.5in]{../Figures/Fig12}\\
  \includegraphics[width=4.5in]{../Figures/Fig13}\\
  \caption[Mass loss versus time for the gasoline and foam fueled fires]{Mass loss versus time for the gasoline and foam fueled fires.}
  \label{Mass}
\end{figure}

@@

Examples of the photographic images are presented in Figs.~\ref{Gas_Foam_Photos} and \ref{Gasoline_Photos}.  Figure 14 is a photograph of the natural gas flame taken during a period of ``steady state'' burning at approximately 80~kW and a photograph of the polyurethane foam fire taken near its peak heat release rate of approximately 60~kW.  Figure 9 presents two photographs of the same gasoline fueled fire when it was burning near its peak heat release rate of 100~kW.  The two photographs, taken 0.2~s apart, represent the range of the observed flame heights from one of the gasoline experiments from 0.65~m and 1.0~m, respectively.

\begin{figure}
  \includegraphics[width=2.7in]{../Figures/Fig14a} 
  \includegraphics[width=2.7in]{../Figures/Fig14b} \\
  \caption[Photographs of the natural gas and polyurethane foam fires]{Photographs showing a natural gas fire (left) and a polyurethane foam fire (right) under the oxygen consumption calorimeter.}
  \label{Gas_Foam_Photos}
\end{figure}

\begin{figure}
  \includegraphics[width=2.7in]{../Figures/Fig15a}
  \includegraphics[width=2.7in]{../Figures/Fig15b} \\
  \caption[Photographs of the gasoline fire]{Photograph showing a gasoline fueled experiment with a flame height of 0.65 m (left) and a photograph taken approximately 0.2 s later with a flame height of 1.0 m (right).}
  \label{Gasoline_Photos}
\end{figure}
